Robotic cloth manipulation for clothing assistance task using Dynamic Movement Primitives

### 2017-02-11 paper/paper.tex
[minor] add v3.0
[major] add v4

\subsection{Robotic cloth manipulation using DMP}
\label{sec:system_overview}
In this section, we provide brief overview of our system. As per the formulation described in section \ref{sec:dmp}, DMP can learn from demonstration. Therefore we start by performing a kinesthetic demonstration with the robot controlled in gravity compensation mode as shown in figure \ref{fig:workflow}. This is referred as ``Teaching Phase'', since in this phase, we are teaching skills to robot to perform the task.  During the demonstration, pose trajectory of end-effector is recorded using Baxter API and stored in a file. The term pose collectively refers to position as cartesian position $p = (p_x, p_y, p_z) \in \mathbb{R}^3$ and orientation as quaternion $q = (q_x, q_y, q_z, q_w) \in \mathbb{R}^4$. Once the demonstration is finished, DMP is parameterized using recorded trajectory file. This is termed as ``Learn Trajectory'' phase. The parameterized DMP can represent all the characteristics of original trajectory. Here, three DMP systems, one for each coordinate axis i.e., $x$, $y$ and $z$ are initialized for one arm. In this way, we have total six DMP systems, which can control both the arms of Baxter robot. In ``Testing Phase'', we change the posture of mannequin by changing the angle of inclination as shown in figure \ref{fig:accuracy}. At this point, we use Kinect Sensor to get the 3D coordinates of wrist and shoulder of mannequin. We change start and goal parameter of DMP trajectories by using this information. In this way, we have generalized DMP system, which can adapt modified posture referred as ``DMP Generalization''. Before using Kinect Sensor, it is extremely important to do Kinect-Baxter calibration, so that 3D coordinates are translated from Kinect space to Baxter space. For Kinect-Baxter calibration, we collect a dataset of points observed by Baxter and Kinect. Then we use absolute orientation calibration \citep{umeyama1991least} to align the frames of reference.

We have divided the complete trajectory into two parts: (a) The reaching part, which refers to the trajectory starts from home position of robot and ends till hands of mannequin (b) the clothing part, which refers to the trajectory starts from hands of mannequin and reaches upto shoulder nearly. The reaching part can be performed through simple position based controller but for the clothing phase, we need to train DMP since this part changes drastically between postures and can be prone to failure.

@@

To evaluate this experiment, we performed it many times and monitored trajectory generated by DMP system. We defined angle of inclination as the angle between arm of mannequin and horizontal axis.  The angle of inclination is defined in clockwise direction, hence it is $+ve$ when arms are inclined upward, similarly it is $-ve$ when arms are inclined downward. It was changed by keeping arms at different-different height. The accuracy measurement is shown in figure \ref{fig:accuracy}. It was observed that DMP system was able to generate the appropriate trajectory for a range of $20^\circ$ and it never failed in this range. However, as we keep on going far away from the original posture, success rate starts declining and finally reached to $0$.

\begin{figure}
	\centering
	\begin{subfigure}{.35\linewidth}
		\includegraphics[width=\linewidth]{inclination}
	\end{subfigure}%
	\begin{subfigure}{.65\linewidth}
		\includegraphics[width=\linewidth]{success_rate}
	\end{subfigure}
	\caption{Accuracy measurement}
	\label{fig:accuracy}
\end{figure}

@@

\section{Discussion}
\label{sec:discussion}
The clothing assistance task deals with manipulation of highly deformable clothing article, which inherits complex dynamics. Failure can be detected by observing end-effector forces. The shape of clothing article keeps on changing during each trail, which affects the task settings. That is why, a trajectory which was able to preform successfully in the past, doesn't work later sometimes. We noticed that there can be many failure scenarios such as clothing article gets stuck into the fingers, sleeve getting stuck on the arms, sleeve not entering the arm but entirely missing etc. Proposed failure detection method by using force information can detect failures when clothing article is stuck tightly. The other reason of failure can be position based DMP system. Orientation plays a major role in the task however DMP system was initialized by position data while ignoring the orientation of end-effector. The orientation was kept same as it was in original trajectory.

>>>>

\subsection{Robotic cloth manipulation using DMP}
\label{sec:system_overview}
In this section, we provide brief overview of our system. As per the formulation described in section \ref{sec:dmp}, DMP can learn from demonstration. Therefore we start by performing a kinesthetic demonstration with the robot controlled in gravity compensation mode as shown in figure \ref{fig:workflow}. This is referred as ``Teaching Phase'', since in this phase, we are teaching skills to robot to perform the task.  During the demonstration, pose trajectory of end-effector is recorded using Baxter API and stored in a file. The term pose collectively refers to position as cartesian position $p = (p_x, p_y, p_z) \in \mathbb{R}^3$ and orientation as quaternion $q = (q_x, q_y, q_z, q_w) \in \mathbb{R}^4$. Once the demonstration is finished, DMP is parameterized using recorded trajectory file. This is termed as ``Learn Trajectory'' phase. The parameterized DMP can represent all the characteristics of original trajectory. Here, three DMP systems, one for each coordinate axis i.e., $x$, $y$ and $z$ are initialized for one arm. In this way, we have total six DMP systems, which can control both the arms of Baxter robot. In ``Testing Phase'', we change the posture of mannequin by changing the angle of inclination as shown in figure \ref{fig:inclination}. At this point, we use Kinect Sensor to get the 3D coordinates of wrist and shoulder of mannequin. We change start and goal parameter of DMP trajectories by using this information. In this way, we have generalized DMP system, which can adapt modified posture referred as ``DMP Generalization''. Before using Kinect Sensor, it is extremely important to do Kinect-Baxter calibration, so that 3D coordinates are translated from Kinect space to Baxter space. For Kinect-Baxter calibration, we collect a dataset of points observed by Baxter and Kinect. Then we use absolute orientation calibration \citep{umeyama1991least} to align the frames of reference.

We have divided the complete trajectory into two parts: (a) The reaching part, which refers to the trajectory starts from home position of robot and ends till hands of mannequin (b) the clothing part, which refers to the trajectory starts from hands of mannequin and reaches up to shoulder nearly. The reaching part can be performed through simple position based controller but for the clothing phase, we need to train DMP since this part changes drastically between postures and can be prone to failure.

@@

To evaluate this experiment, we performed it many times and monitored trajectory generated by DMP system. We defined angle of inclination as the angle between arm of mannequin and horizontal axis. The angle of inclination is defined in clockwise direction, hence it is +ve when arms are inclined upward, similarly it is -ve when arms are inclined downward as shown in figure \ref{fig:inclination}. It was changed by keeping arms at various heights. The angle of inclination for horizontal  orientation is refereed as 0 degree. For each angle of inclination, we performed the experiment 6 times. We then changed the angle of inclination by 5 degree step size and followed the same steps as described above. We stopped this experiment, once we noticed a huge degradation in success rate, such as success rate close to zero. The accuracy measurement is shown in figure \ref{fig:accuracy}. It was observed that DMP system was able to generate the appropriate trajectory for a range of $20^\circ$ and it never failed in this range. However, as we keep on going far away from the original posture, success rate starts declining and finally reached to $0$.

\begin{figure}
	\centering
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=.9\linewidth]{inclination_plus}
	    \caption{Positive angle}
	\end{subfigure}%
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=.9\linewidth]{inclination_minus}
	    \caption{Negative angle}
	\end{subfigure}
	\caption{Angle of Inclination}
	\label{fig:inclination}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{success_rate}
	\caption{Accuracy measurement}
	\label{fig:accuracy}
\end{figure}

@@

\section{Discussion}
\label{sec:discussion}
The clothing assistance task deals with manipulation of highly deformable clothing article, which inherits complex dynamics. Failure can be detected by observing end-effector forces. The shape of clothing article keeps on changing during each trail, which affects the task settings. That is why, a trajectory which was able to preform successfully in the past, does not work later sometimes. We noticed that there can be many failure scenarios such as clothing article gets stuck into the fingers, sleeve getting stuck on the arms, sleeve not entering the arm but entirely missing etc. Proposed failure detection method by using force information can detect failures when clothing article is stuck tightly. The other reason of failure can be position based DMP system. Orientation plays a major role in the task however DMP system was initialized by position data while ignoring the orientation of end-effector. The orientation was kept same as it was in original trajectory.
